---
title: "Results overview"
author: "Andrzej Wójtowicz, Patryk Żywica"
geometry: margin=0.1in
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
classoption: landscape,a4paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, comment="", 
                      echo=FALSE, warning=FALSE, message=FALSE)
```

Document generation date: `r Sys.time()`

# Executive summary

This document presents an overview of the results obtained by the aggregation strategies 
in medical diagnosis support under data incompleteness.

# Introduction

In the *Analytic datasets construction* document we described how to obtain the training and test
sets for the evaluation process. In the **Evaluating aggregation strategies** document we
described how to evaluate different aggregation operators and thresholding strategies on these data.

In the following document we present an overview of the performance of aggregation strategies. To decode names of the aggregation strategies, please refer to comments and code in `aggregators.R` and `aggregators-helpers.R` scripts.

```{r load-data-and-scripts}
load("datasets/evaluation-output.RData")
source("results-overview.R")
```

# Results

This section presents the results obtained on the training and test set. More detailed
resutls consideres selected aggregation strategies which may fulfil medical requirements
in the diagnostic process.

## Training set

The following figure presents the performance top 5 aggregation strategies within each group on the training set (by the lowest average total cost). For the reference, the original and uncertaintified models are also plotted.

```{r plots-training}
agg.sub = subset(binded.stats.all.perf,
                 Measure==PERFORMANCE.MEASURE & Class=="Aggregation") %>%
          group_by(Class, Subclass, Subsubclass)

bestAggregators = NA
if (PERFORMANCE.MEASURE.DESC) {
    bestAggregators = top_n(agg.sub, 5, Value.training)$Method
} else {
    bestAggregators = top_n(agg.sub, 5, -Value.training)$Method }

plotStatsTrainingSet(subset(training.stats.all,
                            Method %in% bestAggregators | Class=="Model"),
                     PERFORMANCE.MEASURE)
```

## Test set

The left part of the following figure compares the total cost performance on the test set among the
original and uncertaintified models and each aggregation group (by the lowest cost). The right 
part compares accuracy, sensitivity, specificity and decisiveness on test set among among the 
original and uncertaintified models and each aggregation group.
 

```{r plots-test}
plotStatsTestSet(test.stats.all,
                  PERFORMANCE.MEASURE, PERFORMANCE.MEASURE.DESC,
                  c("Sensitivity", "Specificity", "Decisiveness", "Accuracy"))
```

## Selected aggregation strategies

The left part of the following figure compares the total cost performance on the test set of the
aggreagation strategies which fulfil medical requirements in the diagnostic process. The right part
compares accuracy, sensitivity, specificity and decisiveness on these aggregation strategies. 

```{r plots-selected}
plotSelectedAggregationOperators(selected.aggrs,
                                 PERFORMANCE.MEASURE, 
                                 c("Sensitivity", "Specificity", "Decisiveness", "Accuracy"))
```

## Difference between aggregation strategies and uncertaintified models

The following table shows the resuls McNemar's test among the selected aggregation
strategies and with relation to the uncertaintified models. The `NaN` values indicate 
that in a given pair of methods classify identically.

```{r mcnemar-pvals}
library(knitr)
library(dplyr)

desc = AGGREGATORS.BINDED.DESCRIPTION %>% 
       filter(Method %in% colnames(pvals))
legend = data.frame(Id=LETTERS[1:length(colnames(pvals))], 
                    Method=colnames(pvals))

kable(join(legend, desc, by="Method"), 
      caption="Legend: short ids for selected aggregation strategies", align='l')
colnames(pvals) = LETTERS[1:length(colnames(pvals))]
rownames(pvals) = gsub("unc. ", "", rownames(pvals))
rownames(pvals) = c(LETTERS[1:length(colnames(pvals))],
                    rownames(pvals)[(length(colnames(pvals))+1):length(rownames(pvals))])

pvals3 = round(pvals,3)
pvals3[is.na(pvals) & !is.nan(pvals)]=""
pvals3 = gsub("^0$", "0.000", pvals3, perl=T)
pvals3 = gsub("^1$", "1.000", pvals3, perl=T)
kable(pvals3, caption="McNemar's test with Benjamini-Hochberg correction among selected aggregation strategies and bettwen the strageties and the uncertaintified models")
```
